\huge{ \emph{Supplement to} Algorithms For Shaping a Particle Swarm\\ With a Shared Control Input Using Boundary Interaction}

### 2016-09-02 RA-L/intro.tex
adding missing refs

This paper provides analytical position control results in two canonical workspaces with frictionless walls \S\ref{subsec:FluidInTank}.  These results are limited in the set of shapes that can be generated.  To extend the range of possible shapes, we introduce wall friction to the system model \S \ref{subsec:WallFriction}.  We prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots \S \ref{sec:PostionControl2Robots}, and extend this proof to prove that a rectangular workspace with high friction boundaries is sufficient to position a swarm of $n$ robots arbitrarily within a subset of the workspace \S \ref{sec:PostionControlnRobots}.
\S \ref{sec:simulation} describes implementations of both position control algorithms in simulation and  \S \ref{sec:expResults} describes experiments with a hardware setup and up to 100 robots, as shown in Fig.~\ref{fig:IntroPic}. After a review of recent related work \S \ref{sec:RelatedWork}, we end with directions for future research \S \ref{sec:conclusion}.
%For controlling $\sigma_{xy}$, we prove that the swarm position covariance $\sigma_{xy}$ is controllable given boundaries with non-zero friction. 
%We then prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position a swarm of $n$ robots. 
%We conclude by designing controllers that efficiently regulate $\sigma_{xy}$.
%This paper
%(1) proves that the swarm position covariance $\sigma_{xy}$ is controllable given boundaries with non-zero friction,  %where do we do this?
%(2) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots, 
%(3) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position a swarm of $n$ robots, 
%(4) shows full-state position control with 2 or more robots using  extensive simulations, and
%(5) demonstrate covariance control on our hardware platform with a large number of hardware robots.
%TODO JOURNAL: design controllers that efficiently regulate $\sigma_{xy}$.
%TODO JOURNAL: We will design Lyapunov-inspired controllers for $\sigma_{xy}$ to prove controllability. 
%TODO JOURNAL:  and rank controllability as a function of friction.
% TODO: JOURNAL: and vary wall friction by laser-cutting boundary walls with a variety of profiles. 
%\begin{figure}[t]
%\centering
%\begin{overpic}[width = \columnwidth]{Covariance.jpg}\end{overpic}
%\vspace{-1em}
%\caption{\label{fig:covFriction} Maintaining group cohesion while steering a swarm through an arbitrary maze requires covariance control.
%}\vspace{-1em}
%\end{figure}
% Our paper is organized as follows.  After a discussion of related work in Section \ref{sec:RelatedWork}, we describe our experimental methods for an online human-user experiment in Section \ref{sec:expMethods}.  We report the results of our experiments in Section \ref{sec:expResults}, discuss the lessons learned in Section \ref{sec:discussion}, and end with concluding remarks in Section \ref{sec:conclusion}.

>>>>

This paper provides analytical position control results in two canonical workspaces with frictionless walls \S\ref{subsec:FluidInTank}.  These results are limited in the set of shapes that can be generated.  To extend the range of possible shapes, we introduce wall friction to the system model \S \ref{subsec:WallFriction}.  We prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots \S \ref{sec:PostionControl2Robots}, and extend this to prove a rectangular workspace with high friction boundaries is sufficient to position a swarm of $n$ robots arbitrarily within a subset of the workspace \S \ref{sec:PostionControlnRobots}.
\S \ref{sec:simulation} describes implementations of both position control algorithms in simulation and  \S \ref{sec:expResults} describes experiments with a hardware setup and up to 100 robots, as shown in Fig.~\ref{fig:IntroPic}. After a review of recent related work \S \ref{sec:RelatedWork}, we end with directions for future research \S \ref{sec:conclusion}.
%For controlling $\sigma_{xy}$, we prove that the swarm position covariance $\sigma_{xy}$ is controllable given boundaries with non-zero friction. 
%We then prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position a swarm of $n$ robots. 
%We conclude by designing controllers that efficiently regulate $\sigma_{xy}$.
%This paper
%(1) proves that the swarm position covariance $\sigma_{xy}$ is controllable given boundaries with non-zero friction,  %where do we do this?
%(2) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots, 
%(3) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position a swarm of $n$ robots, 
%(4) shows full-state position control with 2 or more robots using  extensive simulations, and
%(5) demonstrate covariance control on our hardware platform with a large number of hardware robots.
%TODO JOURNAL: design controllers that efficiently regulate $\sigma_{xy}$.
%TODO JOURNAL: We will design Lyapunov-inspired controllers for $\sigma_{xy}$ to prove controllability. 
%TODO JOURNAL:  and rank controllability as a function of friction.
% TODO: JOURNAL: and vary wall friction by laser-cutting boundary walls with a variety of profiles. 
%\begin{figure}[t]
%\centering
%\begin{overpic}[width = \columnwidth]{Covariance.jpg}\end{overpic}
%\vspace{-1em}
%\caption{\label{fig:covFriction} Maintaining group cohesion while steering a swarm through an arbitrary maze requires covariance control.
%}\vspace{-1em}
%\end{figure}
% Our paper is organized as follows.  After a discussion of related work in Section \ref{sec:RelatedWork}, we describe our experimental methods for an online human-user experiment in Section \ref{sec:expMethods}.  We report the results of our experiments in Section \ref{sec:expResults}, discuss the lessons learned in Section \ref{sec:discussion}, and end with concluding remarks in Section \ref{sec:conclusion}.

### 2016-09-02 RA-L/theory.tex
adding missing refs

\subsection{Using Boundaries: Fluid Settling In a Tank}\label{subsec:FluidInTank}
One method to control a swarm's shape in a bounded workspace is to simply push in a given direction until the swarm conforms to the boundary.
\paragraph{Square workplace}
This section examines the mean, variance,  covariance and correlation of a very large swarm of robots as they move inside a square workplace under the influence of gravity pointing in the direction $\beta$. The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$. Under a global input such as gravity, they flow like water, moving to a side of the workplace and forming a polygonal shape, as shown in Fig.~\ref{fig:SquareFill}.

The range of possible angles for the global input angle $\beta $ is [0,2$\pi $). In this range of angles, the swarm assumes eight different polygonal shapes. The shapes alternate between triangles and trapezoids when the area $A$$<$1/2, and alternate between squares with one corner removed and trapezoids when $A$$>$1/2.

Computing the means, variances, covariance, and correlation requires integrating over the region $R$ containing the swarm.  %This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else. 
The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: %, integrated over the unit square with $x$ and $y$ from 0 to 1:

\begin{align}
\bar{x} &=\frac{\iint_R x \,dx\,dy}{A} \label{eq:meanInSquareWorkspace}
\text{, } \bar{y}=\frac{\iint_R y \,dx\,dy}{A}&\\
%\end{equation}
%\begin{align}
\sigma^2_x &=\frac{\iint_R \left(x-\bar{x}\right)^2  \,dx \,dy}{A}  \label{eq:varInSquareWorkspace}
\text{, } \sigma^2_y =\frac{\iint_R  \left(y-\bar{y}\right)^2 \,dx \,dy}{A}&\\
%\end{align}
%\begin{align}
\sigma_{xy} &= \frac{\iint_R  \left(x-\bar{x}_x\right) \left(y-\bar{y}\right) \, dx \,dy}{A} \label{eq:covAndcorrInSquareWorkspace}
\text{, }\rho_{xy} = \frac{\sigma^2_x}{\sigma_x\sigma_y}&
\end{align}

>>>>

\subsection{Using Boundaries: Fluid Settling In a Tank}\label{subsec:FluidInTank}
One method to control a swarm's shape in a bounded workspace is to simply push in a given direction until the swarm conforms to the boundary.
\paragraph{Square workplace}
%The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: 
This section examines the mean $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ of a very large swarm of robots as they move inside a square workplace under the influence of gravity pointing in the direction $\beta$. The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$. Under a global input such as gravity, they flow like water, moving to a side of the workplace and forming a polygonal shape, as shown in Fig.~\ref{fig:SquareFill}.

The range for the global input angle $\beta $ is [0,2$\pi $). In this range, the swarm assumes eight different polygonal shapes. The shapes alternate between triangles and trapezoids when the area $A$$<$1/2, and alternate between squares with one corner removed and trapezoids when $A$$>$1/2.

Computing the means, variances, covariance, and correlation requires integrating over the region $R$ containing the swarm:  %This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else. 
%The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: %, integrated over the unit square with $x$ and $y$ from 0 to 1:

\begin{align}
\bar{x} &=\frac{\iint_R x \,dx\,dy}{A} \label{eq:meanInSquareWorkspace}
\text{, }\qquad \bar{y}=\frac{\iint_R y \,dx\,dy}{A}
\end{align}
\begin{align}
\sigma^2_x &=\frac{\iint_R \left(x-\bar{x}\right)^2  \,dx \,dy}{A}  \label{eq:varInSquareWorkspace}
\text{, } \sigma^2_y =\frac{\iint_R  \left(y-\bar{y}\right)^2 \,dx \,dy}{A}
\end{align}
\begin{align}
\sigma_{xy} &= \frac{\iint_R  \left(x-\bar{x}_x\right) \left(y-\bar{y}\right) \, dx \,dy}{A} \label{eq:covAndcorrInSquareWorkspace}
\text{, }\rho_{xy} = \frac{\sigma^2_x}{\sigma_x\sigma_y}
\end{align}
